---
title: "Agriculture in Pakistan"
output:
    pdf_document: default
    html_document: default
---

```{r, echo=FALSE, message=FALSE, results='hide'}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

library(tidyverse)
library(patchwork)
library(ggthemes)
library(knitr)
library(broom)
library(haven)

base_year = 2019

df_full <- read_stata("data/DEV309_FullDataset_2019.dta")

# Global data
yields_global   <- read_csv("data/yields-global.csv")
volumes_global  <- read_csv("data/volumes-global.csv")

values_global   <- read_csv("data/values-global.csv")

production_all  <- read_csv("data/production-crops-all-normalized.csv") # this makes yields and volumes all redundant

country_codes <- read_csv("data/wikipedia-iso-country-codes.csv")

cc_map <- country_codes %>%
    select(`Alpha-3 code`, `English short name lower case`, `Numeric code`)

production_global <- production_all %>% 
    select(Year, Area, `Area Code (M49)`, `Item Code`, Item, Element, Unit, Value) %>%
    filter(Unit %in% c("ha", "hg/ha", "tonnes"))

values_global_minimal <- values_global %>%
    select(Year, `Area Code (M49)`, Area, `Item Code` = `Item Code (FAO)`, Item, Element, Unit, Value)

global_agri_data <- rbind(production_global, values_global_minimal) %>%
    mutate(
           `Area Code` = str_replace(`Area Code (M49)`, "'", "")
    ) %>%
    select(-`Area Code (M49)`) %>%
    left_join(cc_map, by=c("Area Code" = "Numeric code")) %>%
    rename(
           "countrycode" = "Alpha-3 code",
           "year" = "Year"
    ) %>%
    select(-"English short name lower case")

global_agri_wide <- global_agri_data %>%
    select(-Unit) %>%
    filter(year >= 1991) %>%
    pivot_wider(names_from="Element", values_from="Value") %>%
    rename(`Gross Production Value` = `Gross Production Value (constant 2014-2016 thousand US$)`) %>%
    mutate(
           `Gross Production Value` = `Gross Production Value` * 1000,
           econ_yield = `Gross Production Value` / `Area harvested`,
           price = `Gross Production Value` / Production
    )

# TODO: better list of comparators
# comparators <- (df_full %>% 
#     filter(regionname == "South Asia") %>% 
#     filter(countrycode != "MDV") %>%
#     select(countrycode) %>% 
#     unique())$countrycode

comparators <- c("PAK", "AFG", "BGD", "BTN", "IND", "LKA", "NPL") 
                 # "VNM", "EGY", "IDN", "MAR", "MMR", "PHL")

```

```{r}
# prepare data
inds <- c(
          "Agri GDP Share"  = "WB_nv_agr_totl_zs",
          "Agri Value Added (current US$)" = "WB_nv_agr_totl_cd",
          "Agri Value Added (constant 2010 US$)" = "WB_nv_agr_totl_kd",
          "GDP per capita (constant 2010 US$)" = "WB_ny_gdp_pcap_kd",

          "Agri Employment" = "WB_sl_agr_empl_zs",
          "Agri Female Employment" = "WB_sl_agr_empl_fe_zs",
          "Agri Male Employment" = "WB_sl_agr_empl_ma_zs",
          "pop-working-age" = "WB_sp_pop_1564_to_zs",
          "population" = "WB_sp_pop_totl",
          "pop-working-age-female" = "WB_sp_pop_1564_fe_in",
          "pop-working-age-male" = "WB_sp_pop_1564_ma_in",
          "pop-working-age2" = "WB_sp_pop_1564_to",

          "Agricultural Land" = "WB_ag_lnd_agri_k2",
          "Arable Land" = "WB_ag_lnd_arbl_ha",
          "Cereal yield" = "WB_ag_yld_crel_kg",
          "Irrigated Land" = "WB_ag_lnd_irig_ag_zs",
          "Irrigated Land2" = "QG_fao_luagrirrac"
        )

cc_prod_data_all <- df_full %>%
    select(year, countrycode, regionname, adminregion, incomelevel, {{inds}}) %>%
    mutate(
           # agri_employ = `Agri Employment` * (`pop-working-age` * population),
           agri_employ = `Agri Employment`/100 * `pop-working-age2`,
           agri_prod = `Agri Value Added (constant 2010 US$)` / agri_employ,
           agri_fem_employ = `Agri Female Employment`/100 * `pop-working-age-female`,
           agri_male_employ = `Agri Male Employment`/100 * `pop-working-age-male`,

           va_per_hectare = `Agri Value Added (constant 2010 US$)` / `Arable Land`,
           va_growth = (`Agri Value Added (constant 2010 US$)` - lag(`Agri Value Added (constant 2010 US$)`) ) / lag(`Agri Value Added (constant 2010 US$)`) 
    )

global_crop_shares <- global_agri_wide %>%
    filter(!is.na(`Area harvested`)) %>%
    filter(!is.na(`Gross Production Value`)) %>%
    group_by(countrycode, year) %>%
    arrange(-`Gross Production Value`) %>%
    mutate(
           prod_share = `Gross Production Value` / sum(`Gross Production Value`),
           area_share = `Area harvested` / sum(`Area harvested`)
    ) %>%
    ungroup()

lag_diff <- 19
global_crop_changes <- global_crop_shares %>%
    group_by(countrycode, Item) %>%
    arrange(year) %>%
    mutate(
           prev_area = lag(`Area harvested`, lag_diff),
           prev_yield = lag(Yield, lag_diff),
           prev_prod = lag(Production, lag_diff),
           prev_prod_val = lag(`Gross Production Value`, lag_diff),
           prev_prod_share = lag(prod_share, lag_diff),
           prev_area_share = lag(area_share, lag_diff),
           prev_price = lag(price, lag_diff),
           prev_econ_yield = lag(econ_yield, lag_diff),

           prod_growth = (Production - prev_prod)/prev_prod * 100,
           prod_val_growth = (`Gross Production Value` - prev_prod_val)/prev_prod_val * 100,
           area_growth = (`Area harvested` - prev_area) / prev_area * 100,
           prod_share_growth = (prod_share - prev_prod_share)/prev_prod_share * 100,
           area_share_growth = (area_share - prev_area_share)/prev_area_share * 100,
           price_growth = (price - prev_price)/prev_price * 100,
           yield_growth = (Yield - prev_yield)/prev_yield * 100,
           econ_yield_growth = (econ_yield - prev_econ_yield) / prev_econ_yield * 100
    ) 

```

Pakistan produces less value added per hectare of Arable land than almost all of South Asia. This metric is also growing slowly. Growth in total value add has been the slowest in the region, and the amount of arable land has not changed. The growth rates in workers, across both male and female, is the highest in the region. Ultimately this results in the lowest value added per worker growth in the region as well. 

We want Pakistan to produce more output per person. To achieve this, we are focusing on producing more food given the land that we have. What are the constraints to this, and can we identify what binds?

```{r fig.height = 12, fig.width=10}

va_hectare_time <- cc_prod_data_all %>%
    filter(regionname == "South Asia") %>%
    filter(countrycode != "MDV") %>%
    ggplot(aes(x=year, y=va_per_hectare, color=countrycode, label=countrycode)) +
    geom_line() +
    geom_text(data =. %>% filter(year == 2016), nudge_x = 2) +
    theme_few() +
    theme(legend.position = "none") + 
    labs(
         title = "Agriculture Value Added per Hectare of Arable Land",
         subtitle = "South Asia",
         y = "",
         x = ""
    )

va_worker_time <- cc_prod_data_all %>%
    filter(regionname == "South Asia") %>%
    filter(!is.na(agri_prod)) %>%
    filter(year == base_year) %>%
    ggplot(aes(x=reorder(countrycode, -agri_prod), y=agri_prod)) +
    geom_bar(stat="identity") + 
    geom_bar(data = . %>% filter(countrycode == "PAK"), fill="orange", stat="identity") + 
    theme_few() +
    labs(
         title = "Value Added per Worker",
         subtitle = paste("Agriculture", base_year),
         y = "",
         x = ""
    )

decompose_hectare_df <- cc_prod_data_all %>%
    group_by(countrycode) %>%
    arrange(year) %>%
    mutate(
           prev_arable_land = lag(`Arable Land`, 16),
           arable_land_change = (`Arable Land` - prev_arable_land)/prev_arable_land * 100,

           prev_va_per_hectare = lag(va_per_hectare, 16),
           va_per_hectare_growth = (va_per_hectare - prev_va_per_hectare)/prev_va_per_hectare * 100
    ) %>%
    ungroup()

growth_hectare <- decompose_hectare_df %>%
    filter(regionname == "South Asia") %>%
    filter(year == 2016) %>%
    filter(!is.na(va_per_hectare_growth)) %>%
    ggplot(aes(x=reorder(countrycode, va_per_hectare_growth), y=va_per_hectare_growth)) +
    geom_bar(stat="identity") +
    geom_bar(data = . %>% filter(countrycode == "PAK"), fill="orange", stat="identity") +
    theme_few() + 
    labs(
         title = paste("Value Added per Hectare"),
         subtitle = paste("Growth between 2016 and", 2016 - 16),
         x = "Country",
         y = ""
    )

n_base_year <- 2000
target_year <- 2019

growth_dat <- cc_prod_data_all %>%
    group_by(countrycode) %>%
    arrange(year) %>%
    filter((year == n_base_year) | (year == target_year)) %>%
    mutate(
        growth = (lead(agri_prod) - agri_prod) / agri_prod * 100
    ) %>%
    filter(year == n_base_year) %>%
    filter(!is.na(growth))

growth_worker <- growth_dat %>%
    filter(regionname == "South Asia") %>%
    ggplot(aes(x=reorder(countrycode, growth), y=growth)) +
    geom_bar(stat="identity") +
    geom_bar(data = . %>% filter(countrycode == "PAK"), fill = "orange", stat="identity") +
    theme_few() + 
    theme(legend.position = "none") + 
    labs(
            title = "Growth in Agri Value Added per Worker" ,
            subtitle = paste(n_base_year, "to", target_year),
            x = "Country",
            y = ""
    )


va_hectare_time / (va_worker_time | growth_worker)

```

What is causing the low growth in value added? Is it because prices are not rising, or because yields are not rising? An analysis shows us that prices are not changing over time. 

We can, however, compare the growth in yields for crops which are important to South Asian countries. This analysis shows us that yields in important crops have not improved in Pakistan. 

```{r fig.height=10, fig.width=10}

global_crop_changes %>%
    filter(countrycode %in% comparators) %>%
    filter(year == base_year) %>%
    group_by(countrycode) %>%
    arrange(-prev_prod_share) %>%
    mutate(position = row_number()) %>%
    ungroup() %>%
    mutate(Item = str_trunc(Item, 10)) %>%
    ggplot(aes(x=position, y=yield_growth)) +
    geom_bar(stat="identity", position="dodge") +
    theme_few() +
    scale_fill_viridis_b() +
    facet_wrap(~countrycode) +
    theme(axis.text.x = element_text(angle = 45, vjust=0.9)) +
    labs(
         title = "Growth in Yield per Crop (KG/Ha)",
         subtitle = paste("Between", base_year - lag_diff, "and", base_year),
         x = "Rank of how important this crop used to be",
         y = "% Growth"
    )

```

We can also see where Pakistans important crops lie relative to median yield and price.

```{r}

global_crop_changes %>%
    filter(countrycode %in% comparators) %>%
    filter(year > 2000) %>%
    # filter(Item %in% crop_changes_table$Item) %>%
    group_by(year, Item) %>%
    mutate(
           rel_yield = Yield / median(Yield, na.rm = T),
           rel_price = price / median(price, na.rm = T)
    ) %>%
    ungroup() %>%
    # arrange(year) %>%
    filter(year == base_year) %>%
    filter(countrycode == "PAK") %>%
    filter(prod_share > 0.005) %>%
    ggplot(aes(x=rel_price, y=rel_yield, color=`prod_share`)) +
    geom_vline(xintercept = 1, linetype = "dashed") +
    geom_hline(yintercept = 1, linetype = "dashed") +
    geom_point(data = . %>% filter(year == base_year), aes(size=`Area harvested`)) +
    geom_text(data = . %>% filter(year == base_year), aes(label=Item), nudge_y = -0.015, nudge_x = 0.025) +
    theme_few() + 
    labs(
         title = "Relative Yield vs Relative Price",
         subtitle = "Comparison vs Comparator countries",
         y = "Yield Relative to Median Yield",
         x = "Price Relative to Median Price"
    )

```

There are no real changes in relative prices. Everything is driven by changes in relative yields. 

```{r}
global_crop_changes %>%
    filter(countrycode %in% comparators) %>%
    group_by(year, Item) %>%
    mutate(
           rel_yield = Yield / median(Yield, na.rm = T),
           rel_price = price / median(price, na.rm = T)
    ) %>%
    ungroup() %>%
    filter(year >= 2000) %>%
    filter(countrycode == "PAK") %>%
    filter(prod_share > 0.01) %>%
    ggplot(aes(x=year, y=rel_yield)) +
    geom_hline(yintercept = 1, linetype = "dashed") +
    geom_line() +
    theme_few() + 
    facet_wrap(~Item, scales="free_y") 
```



